---
title: "AKPopMap"
author: "Courtney Hart"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(scales)
library(ggmap)
library(leaflet)
```


#Prepare Regions Shapefile

```{r}
ak_regions <- read_sf("data/Shapefiles/ak_regions_simp.shp")

plot(ak_regions) # something wrong with peninsula
st_crs(ak_regions) #checks out what projection is being used, it's unprojected 
class(ak_regions) # lists features class
```
An SF obejct is a data frame and work well with TidyVerse. We can also use pipes

coordinate transformations
```{r}
ak_regions_3338 <- ak_regions %>%
  st_transform(crs = 3338)

plot(ak_regions_3338)
```

```{r}
summary(ak_regions_3338)

```

Geometry column is sticky and will stay 'connected' to the othere features in the dataframe
```{r}
ak_regions_3338 %>% 
  select(region)
```


#Prepare Population data

```{r}
pop <- read.csv("data/Shapefiles/alaska_population.csv",
                stringsAsFactors = F)

class(pop)
```
Coerce to sf object using 'st_as_sf'

```{r}
pop_4326 <- st_as_sf(pop,
                     coords = c("lng", "lat"),  # this is what coords are removed 
                                                # lat/long colums and set to geometry
                     crs = 4326, remove = F) #this is what CRS we HAVE, not what we have

st_crs(pop_4326)
head(pop_4326)
```

dataframes are still in different coordinate systems, convert pop to 3338
```{r}
pop_3338 <- pop_4326 %>% 
  st_transform(crs = 3338)
```


# Calculate population by region

```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
```
we picked st_within becasue we want t know which points is within each region
region is polygons and our population data frame is points
    st_intersects - if you joined two polygon dataframes

###calculate total population by region
Geometry contains all the communities within each region...we don't want that. We just need a dataframe that is similar to the beginning where population is associated with a region, instead of the points. Therefore we have to drop the geometry column. (See seond line of next chunk)
    
```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
```

```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")

plot(pop_region_3338)
```

how to use groupby and summarise to combing and summarize across geometries
this dissolves boundaries of other factors
```{r}
pop_mgmt <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop = sum(total_pop))
  # summarise(total_pop = sum(total_pop), do_union = F) - this will keep the other boundaries, not dissolve
plot(pop_mgmt["total_pop"])
```


# Make Maps

```{r}
rivers_3338 <- read_sf("data/Shapefiles/ak_rivers_simp.shp")
st_crs(rivers_3338)

```

we are going to match a bunch of dataframes: river-lines, community points, etc
```{r}
ggplot() +
  geom_sf(data = pop_region_3338, mapping = aes(fill = total_pop)) +
  geom_sf(data = pop_3338, mapping = aes(), size = 0.5) +
  geom_sf(data = rivers_3338, mapping = aes(size = StrOrder), color = "black") +
  scale_size(range = c(0.01, 0.2), guide = F) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "skyblue4", high = "lightsalmon", labels = comma) #general custom color gradient
  
```


# Incorporate base maps using ggmap

Transform/reproject our data
```{r}
pop_3857 <- pop_3338 %>% 
  st_transform(crs = 3857)
```

```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

```{r}
bbox <- c(-170, 52, -130, 70) #defines area you are focusing on
ak_map <- get_stamenmap(bbox, zoom = 4)
ak_map_3858 <- ggmap_bbox_to_3857(ak_map)
```

```{r}
ggmap(ak_map_3858) +
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low = "skyblue4", high = "lightsalmon", labels = comma)
```


# Leaflet

```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

```{r}
pop_region_4326 <- pop_region_3338 %>% 
  st_transform(crs = 4326)
```

```{r}
pal <- colorNumeric(palette = "Blues", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1) %>% 
        addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```

