---
title: "Untitled"
author: "Courtney Hart"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library Chunk - there are warnings associated with these libraries but we set warning = F & message = F so that they don't show up. It is always important to read them first. 
```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
```

# Read and Clean Data

Read in Data - url must be in quotes
```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", 
                           stringsAsFactors = FALSE)

head(catch_original)
```

Remove 'all' column and 'notes' column from original data becasue we won't be doing any analysis on them. 
```{r}
catch_data <- catch_original %>% 
  # select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) - too tedious, remove data you don't want instead of choosing hte column you do want.
  select(-All, -notesRegCode)

head(catch_data)
```

Piping: %<% (control-shift-M)
Instead of using 2 operations:
  df_filtered <- filter(df, ...)
  df_selected <- select(df_filtered, ...)
Use the pipe operator:
  df_cleaned <- df %>% 
              filter(...) %>%
              select(...)
This is more efficient beacuse R doesn't have to write to memory all the intermediate steps, makes code more readible, 

Review Data to see if it is reasonable - look at chinook column
```{r}
summary(catch_data)
```

## Fix Chinook Column

Use 'mutate' function to fix chinook column

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

summary(catch_clean)
```

Investigate where NAs come from - Find that row
```{r}
i <- which(is.na(catch_clean$Chinook))
i
  # there is one row [401] that has a problem
```

look at the 401 row - check out above chunks for changes
```{r}
catch_original[i, ]
```

# Reshape Data

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year),
               names_to = "species", 
               values_to = "count")
head(catch_long)
```

# Summarize Data

# Joins to Region table

